---
title: "IsraelGLM"
author: "Israel Golden"
date: "4/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up the workspace
```{r setup, echo = FALSE}
getwd()
setwd('..')
getwd()

# Load your packages
library(tidyverse)
library(ggplot2)

# Set your ggplot theme
mytheme <- theme_light() +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "right")
theme_set(mytheme)
```

## Load in the data

You can also embed plots, for example:

```{r data}
treedata <- read.csv("/Users/israelgolden/Desktop/School/MEM/Semester 4/ENV 872/GitHub Repos/GoldenGriffithsKnierMalinowski_ENV872_EDA_FinalProject/Data/Processed/Cleaned_Gabon_Forest_Plot_Data.csv", stringsAsFactors = TRUE)

head(treedata)
```

```{r data preparation}
# derive basal area from tree DBH in terms of square cm
treedata <- mutate(treedata, basal_area_cm = (((DBH_mm^2)*pi)/4)/100)

#summarize basal area by Project Plot
#Basal area will be in terms of square meters
SiteSummaries <- treedata %>%
  group_by(Project.Plot) %>%
  summarise(Mean_dist = mean(Distance_km),
            total_basal_area = sum(basal_area_cm),
            StDev_basal_area = sd(basal_area_cm),
            total_species = n_distinct(ID)) %>%
  mutate(total_basal_area = (total_basal_area/10000)) %>%
  mutate(basal_area_per_ha = (total_basal_area/0.25))

# Investigate basal area by species at each site
# DF_5A
DF_5A_basalarea <- treedata %>%
  filter(treedata$Project.Plot == 'DF_5A') %>%
  group_by(ID) %>%
  summarise(species_basal_area = sum(basal_area_cm),
            number_of_trees = n(),
            mean_basal_area = mean(basal_area_cm)) %>%
  mutate(plot_id = 'DF_5A')


#DF_5B
DF_5B_basalarea <- treedata %>%
  filter(treedata$Project.Plot == 'DF_5B') %>%
  group_by(ID) %>%
  summarise(species_basal_area = sum(basal_area_cm),
            number_of_trees = n(),
            mean_basal_area = mean(basal_area_cm)) %>%
  mutate(plot_id = 'DF_5B')
  
#DF_3B
DF_3B_basalarea <- treedata %>%
  filter(treedata$Project.Plot == 'DF_3B') %>%
  group_by(ID) %>%
  summarise(species_basal_area = sum(basal_area_cm),
            number_of_trees = n(),
            mean_basal_area = mean(basal_area_cm)) %>%
  mutate(plot_id = 'DF_3B')

#DF_6A
DF_6A_basalarea <- treedata %>%
  filter(treedata$Project.Plot == 'DF_6A') %>%
  group_by(ID) %>%
  summarise(species_basal_area = sum(basal_area_cm),
            number_of_trees = n(),
            mean_basal_area = mean(basal_area_cm)) %>%
  mutate(plot_id = 'DF_6A')

#DF_6B
DF_6B_basalarea <- treedata %>%
  filter(treedata$Project.Plot == 'DF_6B') %>%
  group_by(ID) %>%
  summarise(species_basal_area = sum(basal_area_cm),
            number_of_trees = n(),
            mean_basal_area = mean(basal_area_cm)) %>%
  mutate(plot_id = 'DF_6B')

#IF_2A
IF_2A_basalarea <- treedata %>%
  filter(treedata$Project.Plot == 'IF_2A') %>%
  group_by(ID) %>%
  summarise(species_basal_area = sum(basal_area_cm),
            number_of_trees = n(),
            mean_basal_area = mean(basal_area_cm)) %>%
  mutate(plot_id = 'IF_2A')

species_by_site_df <- rbind(DF_3B_basalarea, DF_5A_basalarea, DF_5B_basalarea, 
                         DF_6A_basalarea, DF_6B_basalarea, IF_2A_basalarea)

library(stringr)

species_by_site_df$genus <- word(species_by_site_df$ID, 1)

```

# Visualize the data

```{r}
# What is the distribution of basal area in square cm for each tree by site?
ggplot(treedata, aes(x = basal_area_cm, fill = Project.Plot)) + 
  geom_histogram(bins = 40) +
  xlim(0,50) +
  facet_wrap(treedata$Project.Plot) +
  labs(title = "Basal area distribution of study plots")
```
So it seems like most sites are pretty not uniform, and small trees are over-represented at each location. Let's compare sites to see which site has the most species.
```{r}
species_by_plot <- ggplot(data=SiteSummaries, aes(x=Project.Plot, y=total_species, fill = Project.Plot)) +
  geom_bar(stat="identity") +
  theme(legend.position="none")

basal_area_by_plot <- ggplot(data=SiteSummaries, aes(x=Project.Plot, y=total_basal_area, fill = Project.Plot)) +
  geom_bar(stat="identity") +
  theme(legend.position="none")

library(cowplot)
plot_grid(species_by_plot, basal_area_by_plot, nrow = 2, align = 'h', rel_heights = c(1.25, 1))

```
So from this we can see that there are no great disparities between sites when it comes to species. At 146, DF_6B has the most species and at 96 DF_5B has the fewest species. Most sites have right around 100. We can also see how each site compares in terms of basal area - i.e., how densely forested each site is. IF_2A has the lowest basal area where DF_5A has the greatest at around 15 square meters per hectare. Given IF_2A's high species richness but low basal area, these data seem to suggest that IF_2A has many small trees but few large ones. Let's continue our exploration by seeing which genuses contribute the most to each site's basal area.

```{r DF_3B}
# Genus basal area at each site
# install.packages("treemapify")
library(treemapify)

# start by creating a new data frame with basal area summaries by genus (basal are in square meters per hectare)
DF_3B_basalarea$genus <- word(DF_3B_basalarea$ID, 1)
DF_3B_genus_summary <- DF_3B_basalarea %>%
  group_by(genus) %>%
  summarise(genus_basal_area = sum(species_basal_area))

# Plot with a treemap showing the area each genus takes up at each site, extrapolated to a hectare
ggplot(DF_3B_genus_summary, aes(area = genus_basal_area, fill = genus_basal_area, label = genus)) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15,
                    grow = TRUE) +
  scale_fill_viridis_c(option = "turbo")
```

```{r DF_5A}
#DF_5A
DF_5A_basalarea$genus <- word(DF_5A_basalarea$ID, 1)
DF_5A_genus_summary <- DF_5A_basalarea %>%
  group_by(genus) %>%
  summarise(genus_basal_area = sum(species_basal_area))

ggplot(DF_5A_genus_summary, aes(area = genus_basal_area, fill = genus_basal_area, label = genus)) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15,
                    grow = TRUE) +
  scale_fill_viridis_c(option = "turbo")
```

```{r DF_5B}
#DF_5B
DF_5B_basalarea$genus <- word(DF_5B_basalarea$ID, 1)
DF_5B_genus_summary <- DF_5B_basalarea %>%
  group_by(genus) %>%
  summarise(genus_basal_area = sum(species_basal_area))

ggplot(DF_5B_genus_summary, aes(area = genus_basal_area, fill = genus_basal_area, label = genus)) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15,
                    grow = TRUE) +
  scale_fill_viridis_c(option = "turbo")
```

```{r DF_6A}
#DF_6A
DF_6A_basalarea$genus <- word(DF_6A_basalarea$ID, 1)
DF_6A_genus_summary <- DF_6A_basalarea %>%
  group_by(genus) %>%
  summarise(genus_basal_area = sum(species_basal_area))

ggplot(DF_6A_genus_summary, aes(area = genus_basal_area, fill = genus_basal_area, label = genus)) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15,
                    grow = TRUE) +
  scale_fill_viridis_c(option = "turbo")
```

```{r DF_6B}
# DF_6B
DF_6B_basalarea$genus <- word(DF_6B_basalarea$ID, 1)
DF_6B_genus_summary <- DF_6B_basalarea %>%
  group_by(genus) %>%
  summarise(genus_basal_area = sum(species_basal_area))

ggplot(DF_6B_genus_summary, aes(area = genus_basal_area, fill = genus_basal_area, label = genus)) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15,
                    grow = TRUE) +
  scale_fill_viridis_c(option = "turbo")
```

```{r IF_2A}
IF_2A_basalarea$genus <- word(IF_2A_basalarea$ID, 1)
IF_2A_genus_summary <- IF_2A_basalarea %>%
  group_by(genus) %>%
  summarise(genus_basal_area = sum(species_basal_area))

ggplot(IF_2A_genus_summary, aes(area = genus_basal_area, fill = genus_basal_area, label = genus)) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15,
                    grow = TRUE) +
  scale_fill_viridis_c(option = "turbo")

```

From these graphs we can see which genera make up the majority of basal area at each site. So how similar are these sites to one another? I wonder. ANOVA?


Let's begin to see if there is any relationship between basal area, species richness, and distance to towns (i.e., along the defaunation gradient)?.

```{r}
# Scatter plot of distance from town and basal area per acre
ggplot(SiteSummaries, aes(x = Mean_dist, y = basal_area_per_ha, color = Project.Plot)) +
  geom_point() +
  labs(title = "Basal area and distance to developed area") + 
       ylab('Mean Basal Area per hectare') +
       xlab("Distance from developed area (km)") +
  geom_smooth(method=lm, se=FALSE, col='red', size=0.5)

```

```{r}
# What about species richness as it relates to distance from towns?
# Scatter plot of distance from town and basal area per acre
ggplot(SiteSummaries, aes(x = Mean_dist, y = total_species, color = Project.Plot)) +
  geom_point() +
  labs(title = "Species richness and distance to developed area") + 
       ylab('Number of unique species') +
       xlab("Distance from developed area (km)") +
  geom_smooth(method=lm, se=FALSE, col='red', size=0.5)

```

Looking at these graphs, it doesn't appear that there's much of an obvious relationship between basal area, species richness, and distance along the defaunation gradient. Still, looks can be deceiving. Let's feed these data into a linear model to see what relationships can be statistically proved. We begin by checking for correlations among variables with a corrplot. 

```{r}
library(corrplot)
SiteSummaryNumbers <- 
  SiteSummaries %>%
  select(Mean_dist:basal_area_per_ha)
corrplot(cor(SiteSummaryNumbers))
```


from this it appears that basal area per hectare is negatively correlated with mean distance at the same time, there appears to be a weak positive correlation between total species and mean distance. Let's build a model and see how well these correlations predict basal area and species richness.

```{r}
# Linear model describing relationship between basal area per hectare as a function of total species and distance to town. 

basal_area_lm <- lm(data = SiteSummaries, basal_area_per_ha ~ Mean_dist + total_species)
step(basal_area_lm)

summary(basal_area_lm)

# Linear model describing relationship between species richness and distance to town
species_richness_lm <- lm(data = SiteSummaries, total_species ~ Mean_dist)
summary(species_richness_lm)
```

In this linear model we use species richness and distance from developed area to predict the basal area per hectare of a given site. The null hypothesis is that there is no relationship between basal area per hectare, species richness, and distance to town. The alternative hypothesis is that either both species richness and/or distance to town will influence basal area per hectare. The results of the model (p = 0.16) indicate that no such significant (p<0.05) relationships exist in this subset of data. However, based on these observations (n = 6), there does appear to be a weak negative relationship (p = 0.09) between distance to town and basal area per hectare. According to the model, with every additional kilometer of distance there is an associated 1.3 square meter decline in basal area of forested land. Once again, these relationships are not considered significant, which means the explanatory variables included in this model are not sufficient to explain observed patterns in stand density along the defaunation gradient. There is no relationship between distance to town and species richness (p = 0.74).
