---
title: "Wrangling & Visualization DBH"
author: "Tasha Griffiths"
date: "4/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How does defaunation affect forest structure? 

How does the DBH Class Distribution respond to the defaunation gradient?

create a graph with X axis DBH, Y axis number of trees, separated by areas with elephants and areas without elephants. 

```{r load data & setup}

#check working directory location
getwd()


#load package for session
library(tidyverse)
library(formatR)
library(ggplot2)

#load dataset
fulldata <- read.csv("./Data/Processed/Cleaned_Gabon_Forest_Plot_Data.csv", 
                     stringsAsFactors = TRUE)

#basic summaries of dataset
class(fulldata)
summary(fulldata)

#dimensions of dataset
dim(fulldata)

#to view dataset
colnames(fulldata)
str(fulldata)

#set theme
mytheme <- theme_classic(base_size = 12) + 
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom")

theme_set(mytheme)

```

## Graphs

These graphs outline that DBH is higher in plots closer to the city. 5A and 5B have higher frequencies of larger trees as compared to site IF_2A which has far more small DBH trees and very very few large DBH trees. However, overall most of the site histograms resemeble each other and are skewed to the smaller DBH side of the scale. This means there are higher frequencies of thinner trees which tends to indicate a younger forest. 

```{r graphs, echo=FALSE}

#basic visualization
DBHbyplot <- ggplot(fulldata, 
                    aes(x = Project.Plot, y = DBH_mm, color = Project.Plot)) +
  geom_point() + xlab("Project Plot") + ylab( "DBH in Milimeter")
print(DBHbyplot)


#checking distance  
ggplot(fulldata, aes(x = Project.Plot, y = Distance_km)) + geom_point()
#order by closest to city - DF_5A, DF_5B, DF_3B, DF_6A, DF_6B, IF_2A

ordereddata <- arrange(fulldata, Distance_km)

#x axis is DBH, Y axis is frequency  - for each plot as a histogram
#wrangle data for each plotsite
DF_5A_Data <- subset(fulldata, Project.Plot== "DF_5A")
DF_5B_Data <- subset(fulldata, Project.Plot== "DF_5B")
DF_3B_Data <- subset(fulldata, Project.Plot== "DF_3B")
DF_6A_Data <- subset(fulldata, Project.Plot== "DF_6A")
DF_6B_Data <- subset(fulldata, Project.Plot== "DF_6B")
IF_2A_Data <- subset(fulldata, Project.Plot== "IF_2A")

#individual historgram full data
plot_DF_5A <- ggplot(DF_5AData, 
                    aes(x = DBH_mm)) +
  geom_histogram() + xlab("DBH in mm") + ylab( "Frequency")
print(plot_DF_5A)

#individual histogram limited DBH scale
plot_DF_5A <- ggplot(DF_5AData, 
                    aes(x = DBH_mm)) +
  geom_histogram() + xlim(0,550) +
  xlab("DBH in mm") + ylab( "Frequency") + ggtitle("DBH of Site DF_5A")
print(plot_DF_5A)


#histogram of all
allsitesbyDBH <- ggplot(fulldata, aes(x = DBH_mm, color = Project.Plot)) +
  geom_histogram() + facet_wrap(vars(Project.Plot)) + #xlim(0,1000) +
  xlab("DBH in mm") + ylab("Frequency") + ggtitle("DBH across Sites")
print(allsitesbyDBH)

allsitesbyDBH_limitedx <- ggplot(fulldata, aes(x = DBH_mm, color = Project.Plot)) +
  geom_histogram() + facet_wrap(vars(Project.Plot)) + xlim(0,150) +
  xlab("DBH in mm") + ylab("Frequency") + ggtitle("DBH across Sites")
print(allsitesbyDBH_limitedx)


#re-order facet_wrap by distance from city
allsitesordered <- ggplot(transform(fulldata, Project.Plot=factor(Project.Plot, levels=c("DF_5A", "DF_5B", "DF_3B", "DF_6A", "DF_6B", "IF_2A")))) + geom_histogram(aes(x = DBH_mm, color = Project.Plot)) + facet_wrap(vars(Project.Plot)) +
  xlab("DBH in mm") + ylab("Frequency") + ggtitle("DBH across Sites")
print(allsitesordered)

allsitesordered_limitedx <- ggplot(transform(fulldata, Project.Plot=factor(Project.Plot, levels=c("DF_5A", "DF_5B", "DF_3B", "DF_6A", "DF_6B", "IF_2A")))) + geom_histogram(aes(x = DBH_mm, color = Project.Plot)) + facet_wrap(vars(Project.Plot)) + xlim(0,150) +
  xlab("DBH in mm") + ylab("Frequency") + ggtitle("DBH across Sites")
print(allsitesordered_limitedx)



#old
#order plots by distance
DBHbydistance <- ggplot(fulldata, 
                    aes(x = Distance_km, y = DBH_mm, color = Project.Plot)) +
  geom_violin() + xlab("Distance from City") + ylab( "DBH in Milimeter")
print(DBHbydistance)
  
#remove outliers
DBHbydistance2 <- ggplot(fulldata, 
                    aes(x = Distance_km, y = DBH_mm, color = Project.Plot)) +
  geom_violin() + ylim(0,200) + 
  xlab("Distance from City") + ylab( "DBH in Milimeter")
print(DBHbydistance2)




#old ones to look at
ggplot(Neonics, aes(x = Neonics$Endpoint)) +
  geom_bar() +
theme(axis.text.x = element_text(angle = 70, hjust = 1))


NTL.LTER.plot <- ggplot(PeterPaul.chem.nutrients, aes(x = tp_ug, y = po4, 
                                                      color = lakename)) + 
  geom_point() + ylim(0, 50) + #add points and limit outliers
  scale_color_manual(values=c("orange", "steelblue3")) + #separate colors for each lake
  xlab("Total Phosphorus") + ylab("Phosphate") + #change axis labels to be more clear 
  geom_smooth(colour="black", method = lm) #adding linear best fit line
print(NTL.LTER.plot)

```


